---
title: "615 HW3"
author: "Shangchen Han"
date: "10/4/2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyr)
library(dplyr)
library(gapminder)
library(knitr)
library(kableExtra)
library(magrittr)
library(ggplot2)
library(AER)
library(tokenizers)
```


*Problem 1*
```{r}
attach(gapminder)
#How many continents are included in the data set?
data <- gapminder
num_cont <-  data %>% count(continent)
num_cont
```
*So, the number of continents is five.*

```{r}
#How many countrys are included? How many countries per continent?
num_coun <- data %>% count(country)
num_coun
num_coun_per_cont <- data %>% group_by(continent) %>% summarise(country %>% unique %>% length)
num_coun_per_cont
```
*There are 142 countries. And there are 52, 25, 33, 30, 2 countries in Africa, Americas, Asia, Europe, and Oceania, respectively.*

```{r}
#Using the gapminder data, produce a report showing the continents in the dataset, total population per continent, and GDP per capita. Be sure that the table is properly labeled and suitable for inclusion in a printed report.

Per <- data %>% group_by(continent) %>% summarise(population_million = sum(pop)/1000000, GDP_million = sum(gdpPercap)/1000000)
kable(cbind(Per), caption = "Total population and total GDP for each continents", align = "c", booktab =T, format = "latex",longtable = F) %>% kable_styling(latex_options = c("HOLD_position"))
```

```{r}
#Produce a well-labeled table that summarizes GDP per capita for the countries in each continent, contrasting the years 1952 and 2007.
Summary_1952 <- data %>% filter(year == 1952)
Summary_2007 <- data %>% filter(year == 2007)
Per_1952 <- Summary_1952 %>% group_by(continent) %>% summarise(Total_GDP_thousand = sum(gdpPercap)/1000, Ave_GDP_thousand = mean(gdpPercap)/1000, Max_GDP_thousand = max(gdpPercap)/1000, Min_GDP_thousand = min(gdpPercap)/1000)
Per_2007 <- Summary_2007 %>% group_by(continent) %>% summarise(Total_GDP_thousand = sum(gdpPercap)/1000, Ave_GDP_thousand = mean(gdpPercap)/1000, Max_GDP_thousand = max(gdpPercap)/1000, Min_GDP_thousand = min(gdpPercap)/1000)
kable(cbind(Per_1952), caption = "Summary GDP per capita for the countries in  each continents in 1952", align = "c", booktab =T, format = "latex",longtable = F) %>% kable_styling(latex_options = c("HOLD_position"))
kable(cbind(Per_2007), caption = "Summary GDP per capita for the countries in each continents in 2007", align = "c", booktab =T, format = "latex",longtable = F) %>% kable_styling(latex_options = c("HOLD_position"))
```

```{r}
#Product a plot that summarizes the same data as the table. There should be two plots per continent.
Total_1952 <- data %>% filter(year==1952)
Total_1952 <- Total_1952 %>% group_by(continent) %>% summarise(Total_GDP_thousand = sum(gdpPercap)/1000)
Total_1952
barplot(as.matrix(Total_1952[,2]),beside = T,legend.text = T,main = "Total GDP per capita for the countries in each continents in 1952",xlab = "Degree",ylab = "continents",names.arg = c("Africa","America","Asia","Europe","Oceania"),ylim = c(0,200))

Total_2007 <- data %>% filter(year==2007)
Total_2007 <- Total_2007 %>% group_by(continent) %>% summarise(Total_GDP_thousand = sum(gdpPercap)/1000)
Total_2007
barplot(as.matrix(Total_2007[,2]),beside = T,legend.text = T,main = "Total GDP per capita for the countries in each continents in 2007",xlab = "Degree",ylab = "continents",names.arg = c("Africa","America","Asia","Europe","Oceania"),ylim = c(0,800))

```

```{r}
#Which countries in the dataset have had periods of negative population growth? Illustrate your answer with a table or plot.
Asian_countries <- data %>% filter(continent == "Asia")

#For Afghanistan:
Afg <- Asian_countries[1:12,]
plot(y=Afg$pop,x=Afg$year,type = "o",xlab = "years" ,ylab = "pop", main = "Total population in Afghanistan from 1952 to 2007")
```
*There was a decrease from 1977 to 1982 in Afghanistan.*

```{r}
#For Cambodia:
Cam <- Asian_countries[37:48,]
plot(y=Cam$pop,x=Cam$year,type = "o",xlab = "years" ,ylab = "pop", main = "Total population in Afghanistan from 1952 to 2007")
```
*There was a decrease from 1972 to 1977 in Cambodia.*

```{r}
#For Kuwait:
Kuw <- Asian_countries[181:192,]
plot(y=Kuw$pop,x=Kuw$year,type = "o",xlab = "years" ,ylab = "pop", main = "Total population in Afghanistan from 1952 to 2007")
```
*There was a decrease from 1987 to 1992 in Kuwait.*

```{r}
#For Lebanon:
Leb <- Asian_countries[193:204,]
plot(y=Leb$pop,x=Leb$year,type = "o",xlab = "years" ,ylab = "pop", main = "Total population in Afghanistan from 1952 to 2007")
```
*There was a decrease from 1977 to 1987 in Lebanon.*

```{r}
#For West Bank and Gaza:
WBG <- Asian_countries[373:384,]
plot(y=WBG$pop,x=WBG$year,type = "o",xlab = "years" ,ylab = "pop", main = "Total population in Afghanistan from 1952 to 2007")
```
*There was a decrease from 1967 to 1972 in Wesr Bank and Gaza.*

```{r}
#Which countries in the dataset have had the highest rate of growth in per capita GDP? Illustrate your answer with a table or plot.
Highest_rate = Summary_1952 %>% mutate(rate = (Summary_2007$gdpPercap-Summary_1952$gdpPercap)/Summary_1952$gdpPercap)
Highest_rate = Highest_rate %>% filter(rate == max(rate))
Highest_rate
```
*The highest rate of growth country in per capita GDP is Equatorial Guinea with 375.6431%.*



*Problem 2*

```{r}
#Product a plot the contracts the frequency of these four combinations. 
data("Fertility")
MM <- Fertility %>% filter(gender1=="male"& gender2=="male")
MF <- Fertility %>% filter(gender1=="male" & gender2=="female")
FF <- Fertility %>% filter(gender1=="female" & gender2=="female")
FM <- Fertility %>% filter(gender1=="female" & gender2=="male")

slices <- c(67799, 63185,60946,62724)
lbls <- c("MM","MF","FF","FM")
pct <- round(slices/sum(slices),3)
lbls <- paste(lbls,pct)
lbls <- paste(lbls,"%",sep = "")
pie(slices,labels = lbls,col = rainbow(length(lbls)),
    main = "Frequency of these four combinations")
```

```{r}
# Are the frequencies different for women in their 20s and wemen who are older than 29?

Fertility_1 <- Fertility %>% filter(age<30)
Fertility_2 <- Fertility %>% filter(age>29)
MM_1 <- Fertility_1 %>% filter(gender1=="male"& gender2=="male")
MF_1 <- Fertility_1 %>% filter(gender1=="male" & gender2=="female")
FF_1 <- Fertility_1 %>% filter(gender1=="female" & gender2=="female")
FM_1 <- Fertility_1 %>% filter(gender1=="female" & gender2=="male")

MM_2 <- Fertility_2 %>% filter(gender1=="male"& gender2=="male")
MF_2 <- Fertility_2 %>% filter(gender1=="male" & gender2=="female")
FF_2 <- Fertility_2 %>% filter(gender1=="female" & gender2=="female")
FM_2 <- Fertility_2 %>% filter(gender1=="female" & gender2=="male")

slices <- c(24505, 22653,22183,22508)
lbls <- c("MM_1","MF_1","FF_1","FM_1")
pct <- round(slices/sum(slices),3)
lbls <- paste(lbls,pct)
lbls <- paste(lbls,"%",sep = "")
pie(slices,labels = lbls,col = rainbow(length(lbls)),
    main = "Frequency of these four combinations with age under 30")

slices <- c(43294, 43294,38763,40216)
lbls <- c("MM_2","MF_2","FF_2","FM_2")
pct <- round(slices/sum(slices),3)
lbls <- paste(lbls,pct)
lbls <- paste(lbls,"%",sep = "")
pie(slices,labels = lbls,col = rainbow(length(lbls)),
    main = "Frequency of these four combinations with age over 30")
```
*The percentage of MM has been decreased, compared under 30 to over 30. The other 3 parts increased.*

```{r}
#Produce a plot that contrasts the frequency of having more than two children by race and ethnicity.

New_Fertility <- Fertility %>% filter(morekids == "yes")

Afam_Y <- New_Fertility %>% filter(afam == "yes" & hispanic == "no" & other == "no")

Hisp_Y <- New_Fertility %>% filter(afam == "no" & hispanic == "yes" & other == "no")

Other_Y <- New_Fertility %>% filter(afam == "no" & hispanic == "no" & other == "yes")

slices <- c(5933, 5555,2581)
lbls <- c("Afam_Y","Hisp_Y","Other_Y")
pct <- round(slices/sum(slices),3)
lbls <- paste(lbls,pct)
lbls <- paste(lbls,"%",sep = "")
pie(slices,labels = lbls,col = rainbow(length(lbls)),
    main = "Percentage of race and ethnicity")


```


*Problem 3*
```{r}
#Use the mtcars and mpg datasets.
str(mpg)
str(mtcars)
```

```{r}
#How many times does the letter “e” occur in mtcars rownames?
row_n <- row.names(mtcars)
letter = sapply(letters, function(x) x<-sum(x==unlist(strsplit(row_n,""))))
letter
```
*There are 25 letter es which occur in mtcars rownames.*

```{r}
#How many cars in mtcars have the brand Merc?
row_n
```
*There are 7 cars which have the brand Merc.*

```{r}
#How many cars in mpg have the brand(“manufacturer” in mpg) Merc?
Merc = mpg %>% count(manufacturer)
Merc
```
*There are 4 mars in mpg that have the brand Merc.*

```{r}
#Contrast the mileage data for Merc cars as reported in mtcars and mpg. Use tables, plots, and a short explaination.
MPG_1 = mpg %>% filter(manufacturer == "mercury")
MTCARS_1 = mtcars[8:14,]
NAME_mtcars = row.names(MTCARS_1)
tbl_mtcars = cbind(NAME_mtcars, MTCARS_1$mpg)
tbl_mpg = cbind(MPG_1$manufacturer, MPG_1$cty, MPG_1$hwy)

kable(tbl_mtcars, digits = 2, align = "c", format = "latex", booktabs=TRUE, ,caption = "Mileage data for Merc cars in mtcars",col.names = c("manufacturer", "mpg" ))
kable(tbl_mpg, digits = 2, align = "c", format = "latex", booktabs=TRUE, ,caption = "Mileage data for Merc cars in mpg", col.names = c("manufacturer",  "cty" ,"hwy")) 


```

*Problem 4*
```{r}
#Draw a sample of 500,000 rows from the babynames data.
library(babynames)
data = babynames
sub_set <- sample(1:1924655,500000,replace = F)
sub_set <- babynames[sub_set,]
sub_set
```

```{r}
#Produce a tabble that displays the five most popular boy names and girl names in the years 1880,1920, 1960, 2000.
M_name <- sub_set %>% filter(sex=="M")
F_name <- sub_set %>% filter(sex=="F")

M_name_1880 <- M_name %>% filter(year==1880)
F_name_1880 <- F_name %>% filter(year==1880)
M_name_1920 <- M_name %>% filter(year==1920)
F_name_1920 <- F_name %>% filter(year==1920)
M_name_1960 <- M_name %>% filter(year==1960)
F_name_1960 <- F_name %>% filter(year==1960)
M_name_2000 <- M_name %>% filter(year==2000)
F_name_2000 <- F_name %>% filter(year==2000)


M_name_1880 <- M_name_1880[with(M_name_1880,order(n)),]
F_name_1880 <- F_name_1880[with(F_name_1880,order(n)),]
M_name_1920 <- M_name_1920[with(M_name_1920,order(n)),]
F_name_1920 <- F_name_1920[with(F_name_1920,order(n)),]
M_name_1960 <- M_name_1960[with(M_name_1960,order(n)),]
F_name_1960 <- F_name_1960[with(F_name_1960,order(n)),]
M_name_2000 <- M_name_2000[with(M_name_2000,order(n)),]
F_name_2000 <- F_name_2000[with(F_name_2000,order(n)),]

M_name_1880 <- tail(M_name_1880,n=5)
F_name_1880 <- tail(F_name_1880,n=5)
M_name_1920 <- tail(M_name_1920,n=5)
F_name_1920 <- tail(F_name_1920,n=5)
M_name_1960 <- tail(M_name_1960,n=5)
F_name_1960 <- tail(F_name_1960,n=5)
M_name_2000 <- tail(M_name_2000,n=5)
F_name_2000 <- tail(F_name_2000,n=5)

M_1880 <- M_name_1880[3]
F_1880 <- F_name_1880[3]
M_1920 <- M_name_1920[3]
F_1920 <- F_name_1920[3]
M_1960 <- M_name_1960[3]
F_1960 <- F_name_1960[3]
M_2000 <- M_name_2000[3]
F_2000 <- F_name_2000[3]


tbl = cbind(M_1880,F_1880,
            M_1920,F_1920,
            M_1960,F_1960,
            M_2000,F_2000)
colnames(tbl) <-   c('Male', 'Female', 
                     "Male", "Female",
                     'Male', 'Female',
                     "Male", "Female")
kable(tbl, digits = 2,align = "c", format = "latex", booktabs=TRUE, ,caption = "The five most popular boy names and girl names in the years 1880,1920, 1960, 2000") %>% kable_styling() %>%
  add_header_above(c("1880"=2, 
                     "1920"=2,
                     "1960"=2,
                   "2000"=2))
```

```{r}
#What names overlap boys and girls?
names = sub_set %>% group_by(name) %>% summarise(lap = length(sex)) %>% filter(lap>1)
names
```
*There are  51367 names that are overlapped.*

```{r}
#What names were used in the 19th century but have not been used in the 21sth century?
name_19 <- sub_set %>% filter(year<1900)
name_21 <- sub_set %>% filter(year>1999)
name_19 <- name_19 %>% count(name)
name_21 <- name_21 %>% count(name)
name_dif <- subset(name_19, !(name %in% name_21))

```
*There were 3612 names used in the 19th century but not in 21th.*

```{r}
#Produce a chart that shows the relative frequency of the names “Donald”, “Hilary”, “Hillary”, “Joe”,“Barrack”, over the years 1880 through 2017.

Frm_1880_2017 = sub_set %>% filter(year >1879 & year <2018) 
Name_1880_2017 = Frm_1880_2017 %>%filter(name == c("Donald", "Hilary", "Hillary", "Joe", "Barrack"))
y = Name_1880_2017 %>% group_by(name) %>% summarise(n = sum(n))
data =  y %>% mutate(frequency = c(84238/sum(n),847/sum(n),2681/sum(n),21876/sum(n)))
Graph = ggplot(data, aes(x = name,y = frequency)) +
geom_bar(stat = "identity")
print(Graph + ggtitle("Frequency of the names -- Donald, Hilary, Hillary, Joe, Barrack"))
```


