---
title: "MT project"
author: "Shangchen Han"
date: "10/21/2019"
output: ioslides_presentation
---

```{r setup, include=FALSE}
library(tidyr)
library(dplyr)
library(gapminder)
library(knitr)
library(kableExtra)
library(magrittr)
library(ggplot2)
library(lubridate)
knitr::opts_chunk$set(echo = FALSE)
```

##
<div class="centered">
The number and proportion of top 10 taxon in Ningaloo region.
</div>

```{r,echo=FALSE}
df <- read.csv(file = "Bar-chart.csv",header = TRUE,sep = ",")
MyData <- read.csv(file = "data1.csv" ,header = TRUE, sep = ",")
Data1 <- unique(MyData$Ecoregion)
num_Taxon <- MyData %>% count(Taxon)
num_Date <- MyData %>% count(SurveyDate)
Total_Ningaloo <- MyData %>% filter(Ecoregion == "Ningaloo")
Total_Broome <- MyData %>% filter(Ecoregion == "Exmouth to Broome")
Ning_Taxon <- Total_Ningaloo %>% count(Taxon)
Broo_Taxon <- Total_Broome %>% count(Taxon)
Ning_Taxon <- Ning_Taxon[with(Ning_Taxon,order(-n)),]
Ning_Taxon_10 <- Ning_Taxon[1:10,]
New_Ning_Taxon_10 <- Ning_Taxon_10 %>% mutate(prop = n/sum(n))
New_Ning_Taxon_10
```
##
<div class="centered">
The number and proportion of top 10 taxon in Exmouth to Broome region.
</div>

```{r,echo=FALSE}
Broo_Taxon <- Broo_Taxon[with(Broo_Taxon,order(-n)),]
Broo_Taxon_10 <- Broo_Taxon[1:10,]
New_Broo_Taxon_10 <- Broo_Taxon_10 %>% mutate(prop = n/sum(n))
New_Broo_Taxon_10
```  
##
#plot

##
<div class="centered">
Pie chart of ggplot in Ningaloo.
</div>

```{r,echo=FALSE}
ggplot(New_Ning_Taxon_10,aes(x = "", y = prop, fill = Taxon))+
  geom_bar(width = 1, stat = "identity", color = "white")+
  coord_polar("y", start = 0)+
  geom_text(aes(y = prop,label = prop), color = "white")+
  labs(title = "Percentage")
```  

##
<div class="centered">
Bar chart of top 10 taxon in Ningaloo.
</div>
```{r,echo=FALSE}
ggplot(New_Ning_Taxon_10,aes(x=Taxon,y=n))+
  geom_bar(stat = "identity")+
  scale_x_discrete(labels = abbreviate)
```  

##
<div class="centered">
Line chart of top 10 taxon in Ningaloo.
</div>
```{r,echo=FALSE}
ggplot(data = New_Ning_Taxon_10, aes(x=Taxon, y= n, group = 1))+
  geom_line()+
  geom_point()+
  scale_x_discrete(labels = abbreviate)
```  

##
<div class="centered">
Pie chart of top 10 taxon in Ningaloo.
</div>

```{r,echo=FALSE}
slices <- New_Ning_Taxon_10$prop
lbls <- c("Ch-sord","Th-lunare","Sc-bili","Th-lutes","Po-molu","Po-vaiuli","Pl-lac","La-dimi","Ch-ple","Da-aruanus")
pct <-round(New_Ning_Taxon_10$prop,3)
lbls <- paste(lbls,pct)
lbls <- paste(lbls,"%",sep = "")
pie(slices,labels = lbls, col = rainbow(length(lbls)),main = "Percentage of Taxon")
```  


##
<div class="centered">
Pie chart of ggplot in Exmouth to Broome.
</div>
```{r,echo=FALSE}
ggplot(New_Broo_Taxon_10,aes(x = "", y = prop, fill = Taxon))+
  geom_bar(width = 1, stat = "identity", color = "white")+
  coord_polar("y", start = 0)+
  geom_text(aes(y = prop,label = prop), color = "white")+
  labs(title = "Percentage")
```  

##
<div class="centered">
Bar chart of top 10 taxon in Exmouth to Broome.
</div>
```{r,echo=FALSE}
ggplot(New_Broo_Taxon_10,aes(x=Taxon,y=n))+
  geom_bar(stat = "identity")+
  scale_x_discrete(labels = abbreviate)
```  

##
<div class="centered">
Line chart of top 10 taxon in Exmouth to Broome.
</div>
```{r,echo=FALSE}
ggplot(data = New_Broo_Taxon_10, aes(x=Taxon, y= n, group = 1))+
  geom_line()+
  geom_point()+
  scale_x_discrete(labels = abbreviate)
```  

##
<div class="centered">
Pie chart of top 10 taxon in Exmouth to Broome.
</div>
```{r,echo=FALSE}
slices <- New_Broo_Taxon_10$prop
lbls <- c("Po-milleri","Th-lunare","La-dimid","Ab-benga","Sc-bilineata","Lu-carp","Ac-gram","St-obre","Sc-gho","Ha-melano")
pct <-round(New_Broo_Taxon_10$prop,3)
lbls <- paste(lbls,pct)
lbls <- paste(lbls,"%",sep = "")
pie(slices,labels = lbls, col = rainbow(length(lbls)),main = "Percentage of Taxon")
```  

##
<div class="centered">
The number of Thalassoma lunare from 2008 to 2019.
</div>
```{r,echo=FALSE}
data_Th_lunare <- MyData %>% filter(Taxon == "Thalassoma lunare")
num_date_Th_lunare <- data_Th_lunare %>% count(SurveyDate)
New_num_date_Th_lunare <- num_date_Th_lunare %>% mutate(year(num_date_Th_lunare$SurveyDate))
names(New_num_date_Th_lunare)[names(New_num_date_Th_lunare) == "year(num_date_Th_lunare$SurveyDate)"] <- "year"
year_Th <- unique(New_num_date_Th_lunare$year)
sum_Th <- aggregate(New_num_date_Th_lunare$n, by=list(New_num_date_Th_lunare$year), FUN = sum)
sum_Th
```  

##
<div class="centered">
Line chart of number of Thalassoma lunare from 2008 to 2019.
</div>
```{r,echo=FALSE}
ggplot(data = sum_Th,aes(x=Group.1,y=x,group=1))+
  geom_line()+
  geom_point()
```  

##
<div class="centered">

</div>
```{r,echo=FALSE}
Data <- read.csv(file = "Data_Th.csv",header = TRUE, sep = ",")
summary(Data)
a <- Data$Ecoregion
b <- Data$Site
c <- Data$SurveyDate
d <- Data$Depth
glm.fit <- glm(Taxon~a+c+d,family = binomial(link = "logit"),data = Data,control = list(maxit=100))
```  

##
<div class="centered">
The result of regression model.
</div>
```{r,echo=FALSE}
glm.fit
```  
